---
title: "Team10_Dashboard"
output: html_document
date: "2024-11-12"
runtime: shiny
---

## Loading Libraries

The following code chunks loads in the libraries used in this RShiny file.

```{r loading in libraries}
# The following library is for the dashbaord
library(shiny)
# The following libraries is for data cleaning
library(tidyverse)
library(dplyr)
library(tidyr)
library(readr)
# The following libraries are for plotting data
library(ggplot2)
library(scales)
library(plotly)
# The following libraries are for creating the dashboard
library(shiny)
library(shinydashboard)
```

## Loading Data

The following code chunk loads in the survey data used for this project.

```{r loading-in-the-data}
data <- read.csv("survey_data.csv", header=TRUE)
recs <- read.csv("recs.csv", header=TRUE)
```

## Clean Data

The following code chunk cleans the survey data.

```{r data-cleaning}
grocery_data <- data %>% rename(
  Store = Q1, 
  Other_Store = Q2, 
  Produce = Q3_1, 
  Dairy = Q3_2, 
  Grain = Q3_3, 
  Baking = Q3_4, 
  Drinks = Q3_5, 
  Protein = Q3_6, 
  Freq_Dineout = Q1.1, 
  Resturant = Q2.1, 
  Other_Resturant = Q3, 
  Koko = Q4_1, 
  WingsOver = Q4_2, 
  Moosewood = Q4_3, 
  ThompsonBleeker = Q4_4, 
  Pokelava = Q4_5, 
  Appetizers = Q5_1, 
  Entrees = Q5_2, 
  Desserts = Q5_3, 
  Beverages = Q5_4)

grocery_data <- grocery_data[-c(1, 2), ] %>%
  select(-(1:11)) %>%
  filter(!is.na(Store) & Store != "")

grocery_data <- grocery_data %>%
  mutate(across(c("Produce", "Dairy", "Grain", "Baking", "Drinks", "Protein", "Appetizers", "Entrees", "Desserts", "Beverages"), ~ ifelse(. == "", 0, .))) %>% 
  mutate(across(c("Produce", "Dairy", "Grain", "Baking", "Drinks", "Protein","Appetizers", "Entrees", "Desserts", "Beverages"), as.numeric))

grocery_data <- grocery_data %>%
  mutate(Weekly_Grocery_Total = rowSums(across(c("Produce", "Dairy", "Grain", "Baking", "Drinks", "Protein")), na.rm = TRUE)) %>%
   mutate(Weekly_Dineout_Total = rowSums(across(c("Appetizers", "Entrees", "Desserts", "Beverages")), na.rm = TRUE)) %>%
  mutate(Weekly_Total = rowSums(across(c("Weekly_Grocery_Total", "Weekly_Dineout_Total")), na.rm = TRUE))

```

## Grocery vs Dining Out Spending

The folloiwng code chunk is one of our key metrics which is grocery vs dining out spending.

```{r spending-bar-chart}
#weekly grocery vs dining out spending
avg_grocery <- mean(grocery_data$Weekly_Grocery_Total)
avg_dining <- mean(grocery_data$Weekly_Dineout_Total)

averages <- data.frame(
  Category = c("Grocery", "Dineout"),
  Average = c(avg_grocery, avg_dining)
)

# Create bar chart
spending_Barchart <- ggplot(averages, aes(x = Category, y = Average, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0("$", format(round(Average, 2), nsmall = 2))), vjust = -0.3, size = 5) +
  labs(y = "Average Spending (in Dollars)") +
  theme_minimal()

spending_Barchart
```

## Top Restuarant and Grocery Store

The following code chunk is one of our key metrics for top restuarant and grocery store spending.

```{r table-top-rest-store}
top_grocery <- grocery_data %>%
  mutate(week = row_number()) %>% 
  pivot_longer(
    cols = c(Produce, Dairy, Grain, Baking, Drinks, Protein), 
    names_to = "Grocery_Category", 
    values_to = "Grocery_Spending"
  ) %>%
  group_by(week) %>%
  arrange(desc(Grocery_Spending)) %>%
  slice_head(n = 3) %>% 
  
  summarise(
    Top_3_Grocery_Categories = paste(Grocery_Category, collapse = ", "),
    Top_3_Grocery_Amounts = paste(Grocery_Spending, collapse = ", ")
  )


top_restaurant <- grocery_data %>%
  mutate(week = row_number()) %>% 
  pivot_longer(
    cols = c(Appetizers, Entrees, Desserts, Beverages), 
    names_to = "Restaurant_Category", 
    values_to = "Restaurant_Spending"
  ) %>%
  group_by(week) %>%
  arrange(desc(Restaurant_Spending)) %>%
  slice_head(n = 3) %>% 
  
  summarise(
    Top_3_Restaurant_Categories = paste(Restaurant_Category, collapse = ", "),
    Top_3_Restaurant_Amounts = paste(Restaurant_Spending, collapse = ", ")
  )

top_three_spending <- grocery_data %>%
  mutate(week = row_number()) %>% 
  left_join(top_grocery, by = "week") %>%
  left_join(top_restaurant, by = "week") %>%
  select(-week)


top_three_spending %>%
  select(Top_3_Grocery_Categories, Top_3_Grocery_Amounts, Top_3_Restaurant_Categories, Top_3_Restaurant_Amounts)

Table_data <- grocery_data %>%
  mutate(
    Store_Name = ifelse(Store == "Other", Other_Store, Store)
  )

# Step 2: Select the top restaurant (with a value of 1) for each week
Table_data <- Table_data %>%
  rowwise() %>%
  mutate(
    Restaurant_Name = case_when(
      Koko == 1 ~ "Koko",
      WingsOver == 1 ~ "WingsOver",
      Moosewood == 1 ~ "Moosewood",
      ThompsonBleeker == 1 ~ "ThompsonBleeker",
      Pokelava == 1 ~ "Pokelava",
      TRUE ~ NA_character_  # Handle cases where no restaurant is selected
    )
  ) %>%
  ungroup()

# Step 3: Create a separate table with just the store and top restaurant for each week
store_restaurant_table <- Table_data %>%
  mutate(week = row_number()) %>% # Add week number as an identifier
  select(week, Store_Name, Restaurant_Name) # Select only relevant columns

# View the resulting table
store_restaurant_table

```

## Monthly Food Spending

The following code chunk is one of our key metrics for monthly food spending line graph.

```{r monthly-food-spending}

lineGraph_Data <- grocery_data |>
  mutate(
    week = row_number(),
    Date = ''
  ) |>
  filter(Store != "")

start_date <- as.Date("2024-11-01")
lineGraph_Data$Date <- start_date + (lineGraph_Data$week - 1)

fig <- plot_ly(lineGraph_Data, 
               x = ~Date, 
               y = ~Weekly_Total, 
               name = 'Daily Total', 
               type = 'scatter', 
               mode = 'lines+markers',
               line = list(color = 'red'),
               marker = list(color = 'blue'))

lineGraph <- fig |>
  layout(
    xaxis = list(
      title = "Date",
      tickformat = "%m/%d/%Y",         
      range = c(as.Date("2024-10-31"), as.Date("2024-11-30"))
    ),
    yaxis = list(
      title = "Total Spent (in USD)",
      tickprefix = "$",
      range = c(0, 600)
    )
  )

```

## Your Budget is X

The following code chunk is one of our key metrics for demonstrating to users that their budget is X amount.

```{r user-budget-X}

user_budget <- sum(grocery_data$Weekly_Total)

```


## Month's Recommendation

The following code chunk is one of our key metrics for demonstrating to users the current's month recommendation.

```{r month-rec}
# create totals of food budgets
grocery_data <- grocery_data %>%
  mutate(Weekly_Grocery_Total = rowSums(across(c("Produce", "Dairy", "Grain", "Baking", "Drinks", "Protein")), na.rm = TRUE)) %>%
   mutate(Weekly_Dineout_Total = rowSums(across(c("Appetizers", "Entrees", "Desserts", "Beverages")), na.rm = TRUE)) %>%
  mutate(Weekly_Total = rowSums(across(c("Weekly_Grocery_Total", "Weekly_Dineout_Total")), na.rm = TRUE))

# Calculate the total weekly food spend for all respondents
total_weekly_food_spend <- sum(grocery_data$Weekly_Total, na.rm = TRUE)
total_weekly_food_spend

# Calculate the total monthly food spend
total_monthly_food_spend <- total_weekly_food_spend * 4
total_monthly_food_spend

# creating thresholds for food recommendations
groc_thres <- recs %>%
  filter(type == "grocery") %>%
  select("place", "price") %>%
  arrange(price)

res_thres <- recs[c(1:5), ] %>%
  select("place", "price") %>%
  arrange(price)

recsdata <- grocery_data %>%
  select("Weekly_Grocery_Total", "Weekly_Dineout_Total")

# function to match dining budgets to restaurants
recsdata$res_rec <- NA

rest_recs <- data.frame(
  price = recs$price,
  place = recs$place
)

# Function to recommend a restaurant based on budget
res_rec <- function(budget, res_thres) {
  for (i in 1:nrow(res_thres)) {
    if (budget < res_thres$price[i]) {
      return(res_thres$place[i])
    }
  }
  return(res_thres$place[nrow(res_thres)])
}

# Apply the function to each `Weekly_Dineout_Total` in `recsdata`
recsdata$res_rec <- sapply(recsdata$Weekly_Dineout_Total, 
                           function(budget) res_rec(budget, rest_recs))

for (i in 1:nrow(recsdata)) {
  if (recsdata$Weekly_Dineout_Total[i] == 0) {
    recsdata$res_rec[i] <- "No Dining Budget"
  }
}

# Load grocery recommendation data
groc_rec_data <- read.csv("Grocery Data - Sheet1.csv")

# Define the grocery recommendation function
groc_rec <- function(budget, groc_thres) {
  # Iterate through rows of the grocery threshold data
  for (i in 1:nrow(groc_thres)) {
    # Use the correct column name "avg_price"
    if (budget < groc_thres$avg_price[i]) {  
      return(groc_thres$Store[i])  # Return the Store name
    }
  }
  # Return the last option if all prices are above the budget
  return(groc_thres$Store[nrow(groc_thres)])
}

# Apply the `groc_rec` function to each Weekly_Grocery_Total in `recsdata`
recsdata$groc_rec <- sapply(recsdata$Weekly_Grocery_Total, function(budget) groc_rec(budget, groc_rec_data))

# Update `groc_rec` for rows with zero grocery budget
recsdata$groc_rec[recsdata$Weekly_Grocery_Total == 0] <- "No Grocery Budget"

# features going into the dashboard

min_grocery_index <- which.min(recsdata$Weekly_Grocery_Total[recsdata$groc_rec != "No Grocery Budget"])
min_grocery_store <- recsdata$groc_rec[min_grocery_index]
min_grocery_store

min_restaurant_index <- which.min(recsdata$Weekly_Dineout_Total[recsdata$res_rec != "No Dining Budget"])
min_restaurant <- recsdata$res_rec[min_restaurant_index]
min_restaurant
```

## Top Grocery And Restuarant

The following code chunk is one of our key metrics for demonstrating to users the current's top grocery and restuarant spending locations.

```{r top-groc-rest-spending}
top_grocery <- grocery_data %>%
  mutate(week = row_number()) %>% 
  pivot_longer(
    cols = c(Produce, Dairy, Grain, Baking, Drinks, Protein), 
    names_to = "Grocery_Category", 
    values_to = "Grocery_Spending"
  ) %>%
  group_by(week) %>%
  arrange(desc(Grocery_Spending)) %>%
  slice_head(n = 3) %>% 
  
  summarise(
    Top_3_Grocery_Categories = paste(Grocery_Category, collapse = ", "),
    Top_3_Grocery_Amounts = paste(Grocery_Spending, collapse = ", ")
  )


top_restaurant <- grocery_data %>%
  mutate(week = row_number()) %>% 
  pivot_longer(
    cols = c(Appetizers, Entrees, Desserts, Beverages), 
    names_to = "Restaurant_Category", 
    values_to = "Restaurant_Spending"
  ) %>%
  group_by(week) %>%
  arrange(desc(Restaurant_Spending)) %>%
  slice_head(n = 3) %>% 
  
  summarise(
    Top_3_Restaurant_Categories = paste(Restaurant_Category, collapse = ", "),
    Top_3_Restaurant_Amounts = paste(Restaurant_Spending, collapse = ", ")
  )

top_three_spending <- grocery_data %>%
  mutate(week = row_number()) %>% 
  left_join(top_grocery, by = "week") %>%
  left_join(top_restaurant, by = "week") %>%
  select(-week)


top_three_spending_data <- top_three_spending %>%
  select(Top_3_Grocery_Categories, Top_3_Grocery_Amounts, Top_3_Restaurant_Categories, Top_3_Restaurant_Amounts)

grocery_data <- grocery_data %>%
  mutate(
    Store_Name = ifelse(Store == "Other", Other_Store, Store)
  )

# Step 2: Select the top restaurant (with a value of 1) for each week
grocery_data <- grocery_data %>%
  rowwise() %>%
  mutate(
    Restaurant_Name = case_when(
      Koko == 1 ~ "Koko",
      WingsOver == 1 ~ "WingsOver",
      Moosewood == 1 ~ "Moosewood",
      ThompsonBleeker == 1 ~ "ThompsonBleeker",
      Pokelava == 1 ~ "Pokelava",
      TRUE ~ NA_character_  # Handle cases where no restaurant is selected
    )
  ) %>%
  ungroup()

# Step 3: Create a separate table with just the store and top restaurant for each week
store_restaurant_table <- grocery_data %>%
  mutate(week = row_number()) %>% # Add week number as an identifier
  select(week, Store_Name, Restaurant_Name) # Select only relevant columns

# View the resulting table
store_rest_data <- store_restaurant_table |>
  group_by(Store_Name, Restaurant_Name) |>
  count() |>
  filter(!is.na(Restaurant_Name),
         Store_Name != "") |>
  arrange(desc(n))

store_rest_vec <- store_rest_data[3, ]

top_three_spending_data <- top_three_spending_data |>
  filter(Top_3_Restaurant_Amounts == "72, 50, 23")
```

## The Dashboard itself

```{r dashbaord}

# Define UI
ui <- dashboardPage(
  dashboardHeader(title = "November Spending"),
  dashboardSidebar(disable = TRUE), 
  dashboardBody(
    fluidRow(
      column(12,
              valueBox(user_budget, 
               "Monthly Spending (November)", 
               icon = icon("credit-card"),
               width = NULL
               )
      )
    ),
    fluidRow(
      div(
        style = "text-align: left; margin-left: 20px;",
        h2("November's Recommended Grocery Store And Restuarant"),
        p("The following are some grocery stores and restaurants that we 
          recommend based on this month’s spending trends. We hope that 
          by shopping at the suggested store and dining at the recommended
          restaurant, 
          you'll not only save money but also enjoy some great experiences in
          Ithaca!")
      )
    ),
    fluidRow(
      column(6, 
             valueBoxOutput("recommended_grocery",
                            width = NULL)
      ),
      column(6, 
             valueBoxOutput("recommended_restaurant",
                            width = NULL)
      )
    ),
    fluidRow(
      box(
        title = "Spending Trend In November",
        width = 12, status = "primary",
        solidHeader = TRUE,
        plotlyOutput("monthly_spending_plot")
      )
    ),
    fluidRow(
      div(
        style = "text-align: left; margin-left: 20px;",
        h2("November Month's Spending Trend")
      )
    ),
    fluidRow(
      box(
        title = "Average Grocery vs Dineout Spending",
        width = 12, status = "primary", solidHeader = TRUE,
        plotOutput("barChart")
      )
    ),
    fluidRow(
      div(
        style = "text-align: left; margin-left: 20px;",
        h2("Top Three Categories Of Spending"),
        p("Here are the top three food categories you’ve spent 
        the most on at your favorite grocery store and restaurant!")
      )
    ),
    fluidRow(
      box(
        title = "Wegmans Spending", 
        width = 6, 
        solidHeader = TRUE, 
        status = "primary", 
        textOutput("wegmans_spending")
      ),
      box(
        title = "Thompson Bleeker Spending", 
        width = 6, 
        solidHeader = TRUE, 
        status = "primary", 
        textOutput("thompson_bleeker_spending")
      )
    ),
    fluidRow(
      div(
        style = "text-align: left; margin-left: 20px;",
        h2("Enter Your Ideal Total Food Budget"),
        p("Here are the recommended store and restaurant for your shopping and
        dining out. We suggest setting a reasonable budget—anything close to $0
        might not quite
        cover a whole month in Ithaca!")
      )
    ),
    fluidRow(
      column(6,
             numericInput("custom_budget", 
                          "Enter your November budget:", 
                          value = 0, 
                          min = 0,
                          max = 100000
             )
      )
    ),
    fluidRow(
      column(6,
             valueBoxOutput("custom_grocery",
                            width = NULL)
      ),
      column(6,
             valueBoxOutput("custom_restaurant",
                            width = NULL)
      )
    )
  )
)

# Define Server
server <- function(input, output) {
  
  # Grocery and restaurant recommendation thresholds
  groc_thres <- data.frame(
    place = c("Aldi's", "Trader Joes", "Walmart", "Tops", "Wegmans", "Green Star", "BJ"),
    price = c(45.64, 57.19, 62.23, 65.35, 67.91, 89.08, 97.61)
  )
  
  res_thres <- data.frame(
    place = c("Pokelava", "Koko", "Wings Over", "Moosewood", "Thompson and Bleecker"),
    price = c(22.05, 26.98, 28.46, 40.25, 49.00)
  )
  
  # Define the recommendation functions
  groc_rec <- function(budget, groc_thres) {
    for (i in 1:nrow(groc_thres)) {
      if (budget < groc_thres$price[i]) {
        return(groc_thres$place[i])
      }
    }
    return(groc_thres$place[nrow(groc_thres)])
  }
  
  res_rec <- function(budget, res_thres) {
    for (i in 1:nrow(res_thres)) {
      if (budget < res_thres$price[i]) {
        return(res_thres$place[i])
      }
    }
    return(res_thres$place[nrow(res_thres)])
  }
  
  # Static recommendations
  output$user_budget_box <- renderValueBox({
    valueBox(user_budget, "November Budget", icon = icon("credit-card"), color = "blue")
  })
  
  output$recommended_grocery <- renderValueBox({
    valueBox("Wegmans", "Recommended Grocery Store", icon = icon("cart-shopping"), color = "green")
  })
  
  output$recommended_restaurant <- renderValueBox({
    valueBox("Moosewood", "Recommended Restaurant", icon = icon("utensils"), color = "purple")
  })
  
  # Validation function for input
  validate_input <- function(value) {
    if (is.null(value) || !is.numeric(value) || value < 0) {
      return(FALSE)
    }
    return(TRUE)
  }
  
  # Custom recommendations based on user input
  output$custom_grocery <- renderValueBox({
    grocery <- groc_rec(input$custom_budget, groc_thres)
    valueBox(grocery, "Recommended Grocery Store", icon = icon("cart-shopping"), color = "green")
  })
  
  output$custom_restaurant <- renderValueBox({
    restaurant <- res_rec(input$custom_budget, res_thres)
    valueBox(restaurant, "Recommended Restaurant", icon = icon("utensils"), color = "purple")
  })
  
  # Custom recommendations based on user input with validation
  output$custom_grocery <- renderValueBox({
    if (!validate_input(input$custom_budget)) {
      valueBox("Error Invalid Input", 
               "Please provide a positive number greater or equal to zero", 
               icon = icon("exclamation-triangle"), color = "red")
    } else {
      grocery <- groc_rec(input$custom_budget, groc_thres)
      valueBox(grocery, "Recommended Grocery Store", icon = icon("cart-shopping"), color = "yellow")
    }
  })
  
  output$custom_restaurant <- renderValueBox({
    if (!validate_input(input$custom_budget)) {
      valueBox("Error Invalid Input", 
               "Please provide a positive number greater or equal to zero", 
               icon = icon("exclamation-triangle"), color = "red")
    } else {
      restaurant <- res_rec(input$custom_budget, res_thres)
      valueBox(restaurant, "Recommended Restaurant", icon = icon("utensils"), color = "orange")
    }
  })
  
  # Line graph rendering
  output$monthly_spending_plot <- renderPlotly({
    lineGraph
  })
  
  # Bar chart rendering
  output$barChart <- renderPlot({spending_Barchart})
  
  # Text outputs
  output$wegmans_spending <- renderText({
    "Grain: $61, Produce: $60, Drinks: $36"
  })
  
  output$thompson_bleeker_spending <- renderText({
    "Appetizers: $72, Desserts: $50, Entrees: $23"
  })
}

shinyApp(ui, server)



```