---
title: "dashboard test"
output: html_document
date: "2024-11-21"
runtime: shiny
---

## Loading Libraries

The following code chunks loads in the libraries used in this RShiny file.

```{r loading in libraries}
# The following library is for the dashbaord
library(shiny)
# The following libraries is for data cleaning
library(tidyverse)
library(dplyr)
library(tidyr)
library(readr)
# The folloiwng libraries are for plotting data
library(ggplot2)
library(scales)
library(plotly)
# The following libraries are for creating the dashboard
library(shiny)
library(shinydashboard)
```

## Loading Data

The following code chunk loads in the survey data used for this project.

```{r loading-in-the-data}
data <- read.csv("survey_data.csv", header=TRUE)
recs <- read.csv("recs.csv", header=TRUE)
```

## Clean Data

The following code chunk cleans the survey data.

```{r data-cleaning}
grocery_data <- data %>% rename(
  Store = Q1, 
  Other_Store = Q2, 
  Produce = Q3_1, 
  Dairy = Q3_2, 
  Grain = Q3_3, 
  Baking = Q3_4, 
  Drinks = Q3_5, 
  Protein = Q3_6, 
  Freq_Dineout = Q1.1, 
  Resturant = Q2.1, 
  Other_Resturant = Q3, 
  Koko = Q4_1, 
  WingsOver = Q4_2, 
  Moosewood = Q4_3, 
  ThompsonBleeker = Q4_4, 
  Pokelava = Q4_5, 
  Appetizers = Q5_1, 
  Entrees = Q5_2, 
  Desserts = Q5_3, 
  Beverages = Q5_4)

grocery_data <- grocery_data[-c(1, 2), ] %>%
  select(-(1:11)) %>%
  filter(!is.na(Store) & Store != "")

grocery_data <- grocery_data %>%
  mutate(across(c("Produce", "Dairy", "Grain", "Baking", "Drinks", "Protein", "Appetizers", "Entrees", "Desserts", "Beverages"), ~ ifelse(. == "", 0, .))) %>% 
  mutate(across(c("Produce", "Dairy", "Grain", "Baking", "Drinks", "Protein","Appetizers", "Entrees", "Desserts", "Beverages"), as.numeric))

grocery_data <- grocery_data %>%
  mutate(Weekly_Grocery_Total = rowSums(across(c("Produce", "Dairy", "Grain", "Baking", "Drinks", "Protein")), na.rm = TRUE)) %>%
   mutate(Weekly_Dineout_Total = rowSums(across(c("Appetizers", "Entrees", "Desserts", "Beverages")), na.rm = TRUE)) %>%
  mutate(Weekly_Total = rowSums(across(c("Weekly_Grocery_Total", "Weekly_Dineout_Total")), na.rm = TRUE))

```

## Grocery vs Dining Out Spending

The folloiwng code chunk is one of our key metrics which is grocery vs dining out spending.

```{r spending-bar-chart}
#weekly grocery vs dining out spending
avg_grocery <- mean(grocery_data$Weekly_Grocery_Total)
avg_dining <- mean(grocery_data$Weekly_Dineout_Total)

averages <- data.frame(
  Category = c("Grocery", "Dineout"),
  Average = c(avg_grocery, avg_dining)
)

# reordering the categories
averages$Category <- factor(averages$Category, levels = c("Grocery", "Dineout"))

# Creating the bar chart
spending_Barchart <- ggplot(averages, aes(x = Category, y = Average, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0("$", format(round(Average, 2), nsmall = 2))), 
            vjust = ifelse(averages$Average < 5, -1, -0.3), 
            size = 5) +
  labs(
    y = "Average Spending (in Dollars)", 
    x = NULL
  ) +
  scale_fill_manual(values = c("Grocery" = "darkseagreen", "Dineout" = "darkslategray2")) + # Use named colors
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_line(color = "gray", size = 0.5)
  )

spending_Barchart
```

## Top Restuarant and Grocery Store

The following code chunk is one of our key metrics for top restuarant and grocery store spending.

```{r table-top-rest-store}
top_grocery <- grocery_data %>%
  mutate(week = row_number()) %>% 
  pivot_longer(
    cols = c(Produce, Dairy, Grain, Baking, Drinks, Protein), 
    names_to = "Grocery_Category", 
    values_to = "Grocery_Spending"
  ) %>%
  group_by(week) %>%
  arrange(desc(Grocery_Spending)) %>%
  slice_head(n = 3) %>% 
  
  summarise(
    Top_3_Grocery_Categories = paste(Grocery_Category, collapse = ", "),
    Top_3_Grocery_Amounts = paste(Grocery_Spending, collapse = ", ")
  )


top_restaurant <- grocery_data %>%
  mutate(week = row_number()) %>% 
  pivot_longer(
    cols = c(Appetizers, Entrees, Desserts, Beverages), 
    names_to = "Restaurant_Category", 
    values_to = "Restaurant_Spending"
  ) %>%
  group_by(week) %>%
  arrange(desc(Restaurant_Spending)) %>%
  slice_head(n = 3) %>% 
  
  summarise(
    Top_3_Restaurant_Categories = paste(Restaurant_Category, collapse = ", "),
    Top_3_Restaurant_Amounts = paste(Restaurant_Spending, collapse = ", ")
  )

top_three_spending <- grocery_data %>%
  mutate(week = row_number()) %>% 
  left_join(top_grocery, by = "week") %>%
  left_join(top_restaurant, by = "week") %>%
  select(-week)


top_three_spending %>%
  select(Top_3_Grocery_Categories, Top_3_Grocery_Amounts, Top_3_Restaurant_Categories, Top_3_Restaurant_Amounts)

Table_data <- grocery_data %>%
  mutate(
    Store_Name = ifelse(Store == "Other", Other_Store, Store)
  )

# Step 2: Select the top restaurant (with a value of 1) for each week
Table_data <- Table_data %>%
  rowwise() %>%
  mutate(
    Restaurant_Name = case_when(
      Koko == 1 ~ "Koko",
      WingsOver == 1 ~ "WingsOver",
      Moosewood == 1 ~ "Moosewood",
      ThompsonBleeker == 1 ~ "ThompsonBleeker",
      Pokelava == 1 ~ "Pokelava",
      TRUE ~ NA_character_  # Handle cases where no restaurant is selected
    )
  ) %>%
  ungroup()

# Step 3: Create a separate table with just the store and top restaurant for each week
store_restaurant_table <- Table_data %>%
  mutate(week = row_number()) %>% # Add week number as an identifier
  select(week, Store_Name, Restaurant_Name) # Select only relevant columns

# View the resulting table
store_restaurant_table

```

## Monthly Food Spending

The following code chunk is one of our key metrics for monthly food spending line graph.

```{r monthly-food-spending}

# Prepare the data
lineGraph_Data <- grocery_data %>%
  mutate(week = row_number()) %>%
  filter(Store != "") %>%
  filter(week %in% c(1, 2, 3, 4)) %>%
  mutate(Dates = as.Date(c("2024-11-09", "2024-11-15", "2024-11-22", "2024-11-29")))

# Generate a sequence of dates with a 7-day interval within the specified range
date_seq <- seq.Date(from = as.Date("2024-11-09"), 
                     to = as.Date("2024-11-29"), 
                     by = "6 days")

# Create the dynamic plot
fig <- plot_ly(lineGraph_Data, 
               x = ~Dates,  # Use Dates for x-axis
               y = ~Weekly_Total, 
               name = 'Daily Total', 
               type = 'scatter', 
               mode = 'lines+markers',
               line = list(color = '#4e614d'),
               marker = list(color = '#1b291a'))

# Customize layout for interactivity
lineGraph <- fig %>%
  layout(
    xaxis = list(
      title = "Dates",
      showgrid = TRUE,
      tickformat = "%m/%d/%Y",
      tickvals = date_seq, # Set the ticks for the x-axis as the generated date sequence
      ticktext = format(date_seq, "%m/%d/%Y"),  # Customize the format of the date labels
      range = c(as.Date("2024-11-09"), as.Date("2024-11-29"))
    ),
    yaxis = list(
      title = "Total Spent (in USD)",
      tickprefix = "$"
    )
  )

# View the plot
lineGraph

```

## Your Budget is X

The following code chunk is one of our key metrics for demonstrating to users that their budget is X amount.

```{r user-budget-X}

user_budget <- sum(grocery_data$Weekly_Total)

```


## Month's Recommendation

The following code chunk is one of our key metrics for demonstrating to users the current's month recommendation.

```{r month-rec}
# create totals of food budgets
grocery_data <- grocery_data %>%
  mutate(Weekly_Grocery_Total = rowSums(across(c("Produce", "Dairy", "Grain", "Baking", "Drinks", "Protein")), na.rm = TRUE)) %>%
   mutate(Weekly_Dineout_Total = rowSums(across(c("Appetizers", "Entrees", "Desserts", "Beverages")), na.rm = TRUE)) %>%
  mutate(Weekly_Total = rowSums(across(c("Weekly_Grocery_Total", "Weekly_Dineout_Total")), na.rm = TRUE))

# Calculate the total weekly food spend for all respondents
total_weekly_food_spend <- sum(grocery_data$Weekly_Total, na.rm = TRUE)
total_weekly_food_spend

# Calculate the total monthly food spend
total_monthly_food_spend <- total_weekly_food_spend * 4
total_monthly_food_spend

# creating thresholds for food recommendations
groc_thres <- recs %>%
  filter(type == "grocery") %>%
  select("place", "price") %>%
  arrange(price)

res_thres <- recs[c(1:5), ] %>%
  select("place", "price") %>%
  arrange(price)

recsdata <- grocery_data %>%
  select("Weekly_Grocery_Total", "Weekly_Dineout_Total")

# function to match dining budgets to restaurants
recsdata$res_rec <- NA

rest_recs <- data.frame(
  price = recs$price,
  place = recs$place
)

# Function to recommend a restaurant based on budget
res_rec <- function(budget, res_thres) {
  for (i in 1:nrow(res_thres)) {
    if (budget < res_thres$price[i]) {
      return(res_thres$place[i])
    }
  }
  return(res_thres$place[nrow(res_thres)])
}

# Apply the function to each `Weekly_Dineout_Total` in `recsdata`
recsdata$res_rec <- sapply(recsdata$Weekly_Dineout_Total, 
                           function(budget) res_rec(budget, rest_recs))

for (i in 1:nrow(recsdata)) {
  if (recsdata$Weekly_Dineout_Total[i] == 0) {
    recsdata$res_rec[i] <- "No Dining Budget"
  }
}

# Load grocery recommendation data
groc_rec_data <- read.csv("Grocery Data - Sheet1.csv")

# Define the grocery recommendation function
groc_rec <- function(budget, groc_thres) {
  # Iterate through rows of the grocery threshold data
  for (i in 1:nrow(groc_thres)) {
    # Use the correct column name "avg_price"
    if (budget < groc_thres$avg_price[i]) {  
      return(groc_thres$Store[i])  # Return the Store name
    }
  }
  # Return the last option if all prices are above the budget
  return(groc_thres$Store[nrow(groc_thres)])
}

# Apply the `groc_rec` function to each Weekly_Grocery_Total in `recsdata`
recsdata$groc_rec <- sapply(recsdata$Weekly_Grocery_Total, function(budget) groc_rec(budget, groc_rec_data))

# Update `groc_rec` for rows with zero grocery budget
recsdata$groc_rec[recsdata$Weekly_Grocery_Total == 0] <- "No Grocery Budget"

# features going into the dashboard

min_grocery_index <- which.min(recsdata$Weekly_Grocery_Total[recsdata$groc_rec != "No Grocery Budget"])
min_grocery_store <- recsdata$groc_rec[min_grocery_index]
min_grocery_store

min_restaurant_index <- which.min(recsdata$Weekly_Dineout_Total[recsdata$res_rec != "No Dining Budget"])
min_restaurant <- recsdata$res_rec[min_restaurant_index]
min_restaurant
```

## Top Grocery And Restuarant

The following code chunk is one of our key metrics for demonstrating to users the current's top grocery and restuarant spending locations.

```{r top-groc-rest-spending}
top_grocery <- grocery_data %>%
  mutate(week = row_number()) %>% 
  pivot_longer(
    cols = c(Produce, Dairy, Grain, Baking, Drinks, Protein), 
    names_to = "Grocery_Category", 
    values_to = "Grocery_Spending"
  ) %>%
  group_by(week) %>%
  arrange(desc(Grocery_Spending)) %>%
  slice_head(n = 3) %>% 
  
  summarise(
    Top_3_Grocery_Categories = paste(Grocery_Category, collapse = ", "),
    Top_3_Grocery_Amounts = paste(Grocery_Spending, collapse = ", ")
  )


top_restaurant <- grocery_data %>%
  mutate(week = row_number()) %>% 
  pivot_longer(
    cols = c(Appetizers, Entrees, Desserts, Beverages), 
    names_to = "Restaurant_Category", 
    values_to = "Restaurant_Spending"
  ) %>%
  group_by(week) %>%
  arrange(desc(Restaurant_Spending)) %>%
  slice_head(n = 3) %>% 
  
  summarise(
    Top_3_Restaurant_Categories = paste(Restaurant_Category, collapse = ", "),
    Top_3_Restaurant_Amounts = paste(Restaurant_Spending, collapse = ", ")
  )

top_three_spending <- grocery_data %>%
  mutate(week = row_number()) %>% 
  left_join(top_grocery, by = "week") %>%
  left_join(top_restaurant, by = "week") %>%
  select(-week)


top_three_spending_data <- top_three_spending %>%
  select(Top_3_Grocery_Categories, Top_3_Grocery_Amounts, Top_3_Restaurant_Categories, Top_3_Restaurant_Amounts)

grocery_data <- grocery_data %>%
  mutate(
    Store_Name = ifelse(Store == "Other", Other_Store, Store)
  )

# Step 2: Select the top restaurant (with a value of 1) for each week
grocery_data <- grocery_data %>%
  rowwise() %>%
  mutate(
    Restaurant_Name = case_when(
      Koko == 1 ~ "Koko",
      WingsOver == 1 ~ "WingsOver",
      Moosewood == 1 ~ "Moosewood",
      ThompsonBleeker == 1 ~ "ThompsonBleeker",
      Pokelava == 1 ~ "Pokelava",
      TRUE ~ NA_character_  # Handle cases where no restaurant is selected
    )
  ) %>%
  ungroup()

# Step 3: Create a separate table with just the store and top restaurant for each week
store_restaurant_table <- grocery_data %>%
  mutate(week = row_number()) %>% # Add week number as an identifier
  select(week, Store_Name, Restaurant_Name) # Select only relevant columns

# View the resulting table
store_rest_data <- store_restaurant_table |>
  group_by(Store_Name, Restaurant_Name) |>
  count() |>
  filter(!is.na(Restaurant_Name),
         Store_Name != "") |>
  arrange(desc(n))

store_rest_vec <- store_rest_data[3, ]

top_three_spending_data <- top_three_spending_data |>
  filter(Top_3_Restaurant_Amounts == "72, 50, 23")
```

##Take four rows as four weeks (a month)
```{r}
nov <- grocery_data |>
    slice(1:4) 
nov_bud <- sum(nov$Weekly_Total)
```

##top three categories of most mentioned grocery store (aldis) and restaurant (pokelava)
```{r}
tops <- grocery_data |>
  filter(Resturant == "Pokelava") |>
  select(17:20) 

avg <- colMeans(tops)

print(avg)

```


## The Dashboard itself

```{r dashbaord}

# Define UI
ui <- dashboardPage(
  
  dashboardHeader(
    title = span("CU Budget", style = "font-weight: bold; color: white;"),
    titleWidth = 250
  ),
  
  dashboardSidebar(disable = TRUE),  # Disable the sidebar
  
  dashboardBody(
    # Apply styling for olive header and graph header
    # tags$head(tags$style(HTML("
    #   .skin-blue .main-header .navbar {
    #     background-color: olive;
    #   }
    #   .skin-blue .main-header .logo {
    #     background-color: olive;
    #   }
    #   .box.box-primary > .box-header {
    #     background-color: olive;
    #     color: white;
    #   }
    # "))),
    
    fluidRow(
      column(12,
        valueBox(
          div("Monthly Spending (November)",  
              style = "font-size: 18px; font-weight: normal;"),
          div(paste0("$", nov_bud), 
              style = "font-size: 32px; font-weight: bold;"),
          icon = icon("credit-card"),
          width = NULL
        )
      )
    ),
    
    # Line Graph Section
    fluidRow(
      box(
        title = "Spending Trend In November",
        width = 12,
        status = "primary",
        solidHeader = TRUE,
        plotlyOutput("lineGraph")
      )
    ),
    
    # Bar Chart Section
    fluidRow(
      box(
        title = "Average Grocery vs Dineout Spending",
        width = 12,
        status = "primary",
        solidHeader = TRUE,
        plotOutput("barChart")
      )
    ),
    
    # Additional content sections
    fluidRow(
      div(
        style = "text-align: left; margin-left: 20px;",
        h2("November Month's Spending Trend")
      )
    ),
    fluidRow(
      div(
        style = "text-align: left; margin-left: 20px;",
        h2("Top Three Categories Of Spending"),
        p("Here are the top three food categories you’ve spent 
        the most on at your favorite grocery store and restaurant!")
      )
    ),
    fluidRow(
      box(
        title = "Aldi's Spending", 
        width = 6, 
        solidHeader = TRUE, 
        status = "primary", 
        textOutput("aldis_spending")
      ),
      box(
        title = "Pokelava Spending", 
        width = 6, 
        solidHeader = TRUE, 
        status = "primary", 
        textOutput("poke_spending")
      )
    ),
    fluidRow(
      div(
        style = "text-align: left; margin-left: 20px;",
        h2("Enter Your Ideal Total Food Budget"),
        p("Here are the recommended store and restaurant for your shopping and
        dining out. We suggest setting a reasonable budget—anything close to $0
        might not quite cover a whole month in Ithaca!")
      )
    ),
    fluidRow(
      column(6, 
             valueBoxOutput("recommended_grocery", width = NULL)),
      column(6, 
             valueBoxOutput("recommended_restaurant", width = NULL))
    ),
    fluidRow(
      column(6,
             numericInput("custom_budget", 
                          "Enter your weekly food budget:", 
                          value = min(grocery_data$Weekly_Total), 
                          min = min(grocery_data$Weekly_Total),
                          max = 100000)
      )
    ),
    fluidRow(
      column(6,
             valueBoxOutput("custom_grocery", width = NULL)),
      column(6,
             valueBoxOutput("custom_restaurant", width = NULL))
    )
  )
)

# Define Server
server <- function(input, output) {
  
  # Grocery and restaurant recommendation thresholds
  groc_thres <- data.frame(
    place = c("Aldi's", "Trader Joes", "Walmart", "Tops", "Wegmans", "Green Star", "BJ"),
    price = c(45.64, 57.19, 62.23, 65.35, 67.91, 89.08, 97.61)
  )
  
  res_thres <- data.frame(
    place = c("Pokelava", "Koko", "Wings Over", "Moosewood", "Thompson and Bleecker"),
    price = c(22.05, 26.98, 28.46, 40.25, 49.00)
  )
  
  # Recommendation functions
  groc_rec <- function(budget, groc_thres) {
    for (i in 1:nrow(groc_thres)) {
      if (budget < groc_thres$price[i]) {
        return(groc_thres$place[i])
      }
    }
    return(groc_thres$place[nrow(groc_thres)])
  }
  
  res_rec <- function(budget, res_thres) {
    for (i in 1:nrow(res_thres)) {
      if (budget < res_thres$price[i]) {
        return(res_thres$place[i])
      }
    }
    return(res_thres$place[nrow(res_thres)])
  }
  
 # Validation function for input
  validate_input <- function(value) {
    if (is.null(value) || !is.numeric(value) || value < 0) {
      return(FALSE)
    }
    return(TRUE)
  }
  
  # Custom recommendations based on user input with validation
  output$custom_grocery <- renderValueBox({
    if (!validate_input(input$custom_budget)) {
      valueBox("Error Invalid Input", 
               "Please provide a positive number greater or equal to zero", 
               icon = icon("exclamation-triangle"), color = "red")
    } else {
      grocery <- groc_rec(input$custom_budget, groc_thres)
      valueBox(grocery, "Recommended Grocery Store", icon = icon("cart-shopping"), color = "olive")
    }
  })
  
  output$custom_restaurant <- renderValueBox({
    if (!validate_input(input$custom_budget)) {
      valueBox("Error Invalid Input", 
               "Please provide a positive number greater or equal to zero", 
               icon = icon("exclamation-triangle"), color = "red")
    } else {
      restaurant <- res_rec(input$custom_budget, res_thres)
      valueBox(restaurant, "Recommended Restaurant", icon = icon("utensils"), color = "teal")
    }
  })
  
  # Line graph rendering
  output$lineGraph <- renderPlotly({ lineGraph })
  
  # Bar chart rendering
  output$barChart <- renderPlot({ spending_Barchart })
  
  # Text outputs
  output$aldis_spending <- renderText({
    "Produce: $18,  Protein: $15,  Grain: $14"
  })
  
  output$poke_spending <- renderText({
    "Entrees: $19, Appetizers: $7, Desserts: $4"
  })
}

# Launch the Shiny app
shinyApp(ui, server)




```